Tropopause Evolution in a Rapidly Intensifying Tropical Cyclone: A Static Stability Budget Analysis

### 2017-12-07 paper.tex
Added to the Introduction and changed the wording in a few places. Revised Model Setup section and created Discussion section.

The numerical simulations were performed using version 19.4 of Cloud Model 1 (CM1) described in \cite{BryanRotunno2009} and available online at http://www2.mmm.ucar.edu/people/bryan/cm1/.
The fully-compressible equations of motion were integrated on a 3000-km-wide, 35-km-deep, axisymmetric grid with 1-km horizontal and 250-m vertical grid spacing.
Horizontal turbulence was parameterized using the Smagorinsky scheme described in \citeauthor{BryanRotunno2009} (\citeyear{BryanRotunno2009}, pg. 1773), with a prescribed mixing length that varied linearly from 100 m at a surface pressure of 1015 hPa to 1000 m at a surface pressure of 900 hPa.
This formulation allows for realistically-large horizontal mixing lengths near the hurricane's inner core, consistent with the results of \cite{Bryan2012}, while not over-representing horizontal turbulence in convection at outer radii.
Vertical turbulence was parameterized using the formulation of \citeauthor{MarkowskiBryan2016} (\citeyear{MarkowskiBryan2016}, their Eq. 6), using an asymptotic vertical mixing length of 100 m.
A Rayleigh damping layer was applied outside of the 2900-km radius and above the 25-km level to prevent spurious gravity wave reflection at the model boundaries.
Microphysical processes were parameterized using the \cite{Thompson} microphysics scheme and radiative heating tendencies were computed every two minutes using the RRTMG longwave and shortwave schemes \citep{Iacono}.
A horizontally-homogeneous temperature and humidity field was initialized with a mean sounding computed using all dropsondes deployed during the TCI flight conducted within and around Tropical Storm Patricia on 21 October, 2015 (see \citeauthor{DoyleTCI} \citeyear{DoyleTCI} for details.)
Above 19 km, where few TCI observations were available, the temperature profile was assumed to be isothermal.
Since relative humidity measurements were unreliable at temperatures below -40\textdegree C \citep{BellTCI}, relative humidity was set equal to 50\% above 11.5 km (the level above which temperature dropped below -40\textdegree C).
The vortex described in \citeauthor{RotunnoEmanuel} (\citeyear{RotunnoEmanuel}, their Eq. 37) was used to initialize the wind field, setting all parameters equal to the values used therein.

@@

The static stability can be expressed as the squared Brunt V{\"a}is{\"a}l{\"a} frequency:
   \begin{equation} \label{eq:n2moist}
   EQ. 18 OF BRYAN AND ROTUNNO. ALSO USE EQ. 19
   \end{equation}
In the tropopause layer, $q_s$, $\frac{\partial q_s}{\partial T}$, and $\frac{\partial q_t}{\partial z}$ approach zero; thus, Eq. \ref{eq:n2moist} reduces to:
   \begin{equation} \label{eq:n2dry}
   N^2 = \frac{g}{\theta_v}\frac{\partial \theta_v}{\partial z},
   \end{equation}
where \textit{g} is gravitational acceleration and $\theta_v$ is the virtual potential temperature.
To compute $N^2$, CM1 uses Eq. \ref{eq:n2moist} in saturated environments and Eq. \ref{eq:n2dry} in unsaturated environments.
For mathematical simplicity, only Eq. \ref{eq:n2dry} will be used for the budget computations herein\footnote{The validity of this approximation will be substantiated later in this section.}.
Taking the time derivative of Eq.\ref{eq:n2dry} yields the static stability tendency:

@@

% text...
% \section{Section title}

>>>>

The numerical simulations were performed using version 19.4 of Cloud Model 1 (CM1) described in \cite{BryanRotunno2009} and available online at http://www2.mmm.ucar.edu/people/bryan/cm1/.
The equations of motion were integrated on a 3000-km-wide, 30-km-deep grid axisymmetric grid with 1-km horizontal and 250-m vertical grid spacing.
The computations were performed on an \textit{f}-plane at 15\textdegree{N} latitude, over a sea surface with constant temperature of 30.5\textdegree C, which matches that observed near Hurricane Patricia (2015; \citeauthor{Kimberlainetal2016} \citeyear{Kimberlainetal2016}.
Horizontal turbulence was parameterized using the Smagorinsky scheme described in \citeauthor{BryanRotunno2009} (\citeyear{BryanRotunno2009}, pg. 1773), with a prescribed mixing length that varied linearly from 100 m at a surface pressure of 1015 hPa to 1000 m at a surface pressure of 900 hPa.
This formulation allows for realistically-large horizontal mixing lengths near the hurricane's inner core, consistent with the results of \cite{Bryan2012}, while not over-representing horizontal turbulence in convection at outer radii.
Vertical turbulence was parameterized using the formulation of \citeauthor{MarkowskiBryan2016} (\citeyear{MarkowskiBryan2016}, their Eq. 6), using an asymptotic vertical mixing length of 100 m.
A Rayleigh damping layer was applied outside of the 2900-km radius and above the 25-km level to prevent spurious gravity wave reflection at the model boundaries.
Microphysical processes were parameterized using the \cite{Thompson} microphysics scheme and radiative heating tendencies were computed every two minutes using the Rapid Radiative Transfer Model WHAT DOES THE G STAND FOR? (RRTMG) longwave and shortwave schemes \citep{Iacono}.
A horizontally-homogeneous temperature and humidity field was initialized with a mean sounding computed using all dropsondes deployed during the TCI flight conducted within and around Tropical Storm Patricia on 21 October, 2015 (see \citeauthor{DoyleTCI} \citeyear{DoyleTCI} for details.)
Above 19 km, where few TCI observations were available, the temperature profile was taken from the Climate Forecast System Reanalysis (CFSR) grid point nearest Patricia's storm center, valid at 18 UTC 21 October, 2015.
Since relative humidity measurements were unreliable at temperatures below -40\textdegree C \citep{BellTCI}, relative humidity was set equal to 50\% above 11.5 km (the level above which temperature dropped below -40\textdegree C).
The vortex described in \citeauthor{RotunnoEmanuel} (\citeyear{RotunnoEmanuel}, their Eq. 37) was used to initialize the wind field, setting all parameters equal to the values used therein.

@@

The static stability can be expressed as the squared Brunt V{\"a}is{\"a}l{\"a} frequency:
   \begin{equation} \label{eq:n2moist}
   EQ. 18 OF BRYAN AND ROTUNNO. ALSO USE EQ. 19
   \end{equation}
In the tropopause layer, $q_s$, $\frac{\partial q_s}{\partial T}$, and $\frac{\partial q_t}{\partial z}$ are small. Assuming that these quantities are zero, Eq. \ref{eq:n2moist} reduces to:
   \begin{equation} \label{eq:n2dry}
   N^2 = \frac{g}{\theta_v}\frac{\partial \theta_v}{\partial z},
   \end{equation}
where \textit{g} is gravitational acceleration and $\theta_v$ is the virtual potential temperature.
To compute $N^2$, CM1 uses Eq. \ref{eq:n2moist} in saturated environments and Eq. \ref{eq:n2dry} in unsaturated environments; for mathematical simplicity, however, only Eq. \ref{eq:n2dry} will be used for the budget computations herein\footnote{The validity of this approximation will be substantiated later in this section.}.

Taking the time derivative of Eq. \ref{eq:n2dry} yields the static stability tendency:

The potential temperature tendency can be partitioned into ### "budget variables," which are output directly by the model every two minutes:

POTENTIAL TEMPERATURE TENDENCY EQUATION

@@

  \section{Discussion}

Dunion et al. speculate that the diurna pulse only occurs in mature storms. Maybe the development of the near-tropopause stable layer could partially explain the reason for this.
% text...
% \section{Section title}
